---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# annotaR

<!-- badges: start -->
<!-- badges: end -->

The goal of `annotaR` is to provide a tidy, pipe-based framework for the integrated annotation of gene lists. It streamlines the process of fetching and combining functional annotations, disease associations, and known drug information from multiple bioinformatics databases.

## Installation

You can install the development version of annotaR from GitHub with:

``` r
# install.packages("devtools")
devtools::install_github("YOUR_GITHUB_USERNAME/annotaR")
```

## Example Workflow

`annotaR` uses a pipe-based (`%>%`) workflow to progressively add layers of information to your initial gene list.

1.  **Start with a list of genes.**
2.  **Add functional annotations** (e.g., Gene Ontology terms from g:Profiler).
3.  **Add disease and drug data** (from OpenTargets).
4.  **Visualize the results.**

Here is a quick example using a small list of cancer-related genes:

```{r example, eval=FALSE}
library(annotaR)
library(dplyr)

# 1. Define genes and initialize pipeline
genes_of_interest <- c("TP53", "EGFR", "BRCA1", "KRAS", "BRAF")
annotaR_obj <- annotaR(genes_of_interest)

# 2. Add annotations in a single pipeline
full_annotation <- annotaR_obj %>%
  add_go_terms(sources = c("GO:BP")) %>%
  add_disease_links() %>%
  add_drug_links()

# 3. Explore the results
# Filter for high-confidence disease links
full_annotation %>%
  filter(association_score > 0.8) %>%
  head()
  
# 4. Create a plot
plot_enrichment_dotplot(full_annotation, n_terms = 15)
```
This workflow generates a rich, tidy data frame containing integrated information, ready for downstream analysis and visualization.
